Tropical Andes Functional diversity for plants and Frugivores

### 2024-06-03 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
remove unnecessary code and add histogram

spatial_fdis_frugivore$cellid <- sapply(1:nrow(TAGrid_100km), function(i) {
  contains <- st_contains(TAGrid_100km[i, ], spatial_fdis_frugivore)
  cellid <- ifelse(any(contains), i, NA)
  return(cellid)
})

@@

```{r}
FDisplot <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis") +
    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12))
FDisplot
```

@@

## Turning the FD values into a raster
Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)

```{r}
# Create an empty raster with desired resolution and extent
resolution <- c(0.75, 0.75)  # Set the desired resolution
extent <- c(-85, -54, -24, 14)  # Set the desired extent
empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
```

```{r}
# Generate coordinates
subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site_final) >= 5, ]
subset_coords_plant_sp <-subset_coords_plant[,1:2 ]

plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)

# Convert the dataframe to sf format
spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
```

```{r}
# Rasterize the sf data to create the FD raster
fd_raster_plant <- rasterize(spatial_fdis_plant, empty_raster)
writeRaster(fd_raster_plant$fdis_plant, filename = file.path(output_path,"FD_plant_TA.tif"), format="GTiff",overwrite=T)
writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")
```

## Project FD objects
```{r}
# set crs of sf objects
spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
```

```{r}
spatial_fdis_plant_projected <- st_transform(spatial_fdis_plant, 5389)
```

```{r}
# Raster to raster point
spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
# Convert to a 'conventional' dataframe
fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
```

```{r}
plantgridFDTA <-
  ggplot() +
  geom_sf(data = Americas, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
plantgridFDTA
```

# Calculating Rao's Q
Calculate Rao's Q using occurrence data and trait dissimilarity

```{r}
## Define different spatial scales (e.g., grid cell sizes)
spatial_scales <- c(1000, 5000, 10000, 25000, 50000, 75000, 100000)  # In meters

```{r}
## Create an empty dataframe to store results
plant_raos_q_results <- data.frame(spatial_scale = numeric(),
                             raos_q = numeric())

## Iterate over different spatial scales
for (scale in spatial_scales) {
  ### Aggregate occurrence data to the current spatial scale
  plant_occurrence_aggregated <- aggregate_occurrence(plant_abundance_per_species, scale)

  ### Calculate dissimilarities using trait data
  plant_trait_dissimilarities <- vegdist(plant_traits, method = "bray")

  ### Calculate dissimilarities using occurrence data
  plant_occurrence_dissimilarities <- vegdist(plant_occurrence_aggregated, method = "bray")

  ### Calculate Rao's Q index
  raos_q <- raosq(plant_trait_dissimilarities, plant_occurrence_dissimilarities)

  #### Add results to the dataframe
  plant_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
}
```

```{r}
## Create an empty dataframe to store results
frugivore_raos_q_results <- data.frame(spatial_scale = numeric(),
                             raos_q = numeric())

## Iterate over different spatial scales
for (scale in spatial_scales) {
  ### Aggregate occurrence data to the current spatial scale
  frugivore_occurrence_aggregated <- aggregate_occurrence(frugivore_abundance_per_species, scale)

  ### Calculate dissimilarities using trait data
  frugivore_trait_dissimilarities <- vegdist(frugivore_traits, method = "bray")

  ### Calculate dissimilarities using occurrence data
  frugivore_occurrence_dissimilarities <- vegdist(frugivore_occurrence_aggregated, method = "bray")

  ### Calculate Rao's Q index
  raos_q <- raosq(frugivore_trait_dissimilarities, frugivore_occurrence_dissimilarities)

  #### Add results to the dataframe
  frugivore_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
}
```

@@

```{r}
plot(spatial_fdis_plant$fdis_plant, spatial_fdis_frugivore$fdis_frugivore,
     main = "Plant FD vs Frugivore FD",
     xlab = "Plant Functional Dispersion by cell", ylab = "Frugivore Functional Dispersion by cell")
abline(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
summary(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
```

>>>>

spatial_fdis_frugivore$cellid <- st_intersects(spatial_fdis_frugivore, TAGrid_100km)$nn

@@

```{r}
frugivoregridFDisTA_100km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis") +
    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12))
frugivoregridFDisTA_100km
```

```{r}
# Extract cell values
frugivore_cellFDis_100km <- spatial_fdis_frugivore_grid
hist(frugivore_cellFDis_100km$fdis_value)
```

@@

@@
